---
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
geometry: margin=2.54cm
title: "Analyzing Trends in Soil Moisture and Temperature at SCAN Sites in California from 2015-2019"
subtitle: "https://github.com/sierrakindley/Kindley_ENV872_Project"
author: "Sierra Kindley"
fontsize: 12pt
mainfont: Times New Roman

---

\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

```{r setup, include=FALSE}
# Set and verify working directory
setwd("/Users/sierr/Documents/Duke University/ENVIRON-872L/Kindley_ENV872_Project")
getwd() 

# Load packages
library(tidyverse) 
library(lubridate) 
library(plyr)
library(dplyr)
library(ggplot2)
library(cowplot) 
library(RColorBrewer)

# Set ggplot theme
sierratheme <- theme_classic(base_size = 14) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "top", 
        plot.title = element_text(face = "bold", size = 16, 
                                  color = "black", hjust = 0.5))
theme_set(sierratheme)

# Load datasets
SCAN.BodieHills.CA <- read.csv("./Data/Raw/SCAN_BodieHillsCA_raw.csv")
SCAN.DeepSprings.CA <- read.csv("./Data/Raw/SCAN_DeepSpringsCA_raw.csv")
SCAN.FrenchGulch.CA <- read.csv("./Data/Raw/SCAN_FrenchGulchCA_raw.csv")
SCAN.CochoraRanch.CA <- read.csv("./Data/Raw/SCAN_CochoraRanchCA_raw.csv")
SCAN.FordDryLake.CA <- read.csv("./Data/Raw/SCAN_FordDryLakeCA_raw.csv")
```

# Rationale and Research Questions
The primary aim of this project is to analyze the daily soil moisture and climate data from the selected sites in California and identify trends in soil moisture and soil temperature relative to soil depth, solar radiation, site (specifically with regards to elevation--all sites are located at different elevations), and time of year. These analyses will be performed on data from 2015 to 2019 (01/01/2015 to 12/31/2019).

\newpage

# Dataset Information
Data were collected using the Soil Climate Analysis Network (SCAN) (https://wcc.sc.egov.usda.gov/nwcc/rgrpt?report=daily_scan_por). SCAN is a comprehensive, nationwide soil moisture and climate information system designed to provide data to support natural resource assessments and conservation activities. Administered by the United States Department of Agriculture Natural Resources Conservation Service (NRCS) through the National Water and Climate Center (NWCC), in cooperation with the NRCS National Soil Survey Center, the system focuses on agricultural areas of the U.S. monitoring soil temperature and soil moisture content at several depths, soil water level, air temperature, relative humidity, solar radiation, wind, precipitation, barometric pressure, and more. 

The following selections were made within the dataset:
* California (Geographic Area)
* Bodie Hills, Cochora Ranch, Deep Springs, Ford Dry Lake, French Gulch (Site Selection)
* Daily SCAN Standard Report - Period of Record (Data Report Type)
* Download CSV (spreadsheet)

The five Soil Climate Analysis Network (SCAN) datasets included in this repository contain daily soil moisture and climate data at sites located in the state of California. Each dataset contains data from a single sampling site in California, so a total of five sites are represented. These sites are located at different elevations and in various regions of the state.

\newpage

# Exploratory Analysis 



\newpage

# Analysis
```{r, echo = FALSE}
#Faceted plot of solar radiation by site   
Plot.SolarRad.AllSites <- 
  ggplot(SCAN.CA.1519.summaries, aes(x = date, y = solar_rad_avg, color = station_name_elev)) +
  geom_point(cex = 2) +
  facet_wrap(vars(station_name_elev), nrow = 5) + #create facets by station
  xlab(expression(paste("Date"))) +
  ylab(expression(paste("Mean Monthly Solar Radiaion (W/m"^2*")"))) +
  labs(title = "Mean Monthly Solar Radiation Across California SCAN Stations") +
  scale_x_date(date_breaks = "1 year", date_labels = c("2020", "2015", "2016", "2017", "2018", "2019")) +
  scale_color_manual(values = c("#CC9966", "#933366", "#CC6699", "#9966CC", "#CC99FF")) +
  theme(legend.position = "none") #remove legend from plot
print(Plot.SolarRad.AllSites)

#Create caption for plot
Caption.SolarRad.AllSites <- 
  ggplot() +
  theme(axis.line = element_blank(), axis.text =
          element_text(color = "black", size = 9),
          plot.title = element_text(size = 9, 
                                  color = "black")) + 
  labs(title = "Figure 1. Mean monthly solar radiation (watts/square meter) accross five Soil Climate Analysis Network (SCAN) stations in California from 2015-2019. Each of the \n five SCAN stations are located at different elevations and in various regions of California. The value listed beside each station name denotes the elevation, in feet \n above sea level, of that given site. Please note: Data was not collected at the French Gulch station from May 2016-April 2017. However, this site was included in \n this analysis, as the trends evident in the present data are valuable to understanding trends at similar elevations/sites over time.")

#Combine plot and caption into single figure
Plot.SolarRad.AllSites.withcaption <-
  cowplot::plot_grid(Plot.SolarRad.AllSites, Caption.SolarRad.AllSites, 
                     align = 'h', axis = "a", 
                     nrow = 2, rel_heights = c(.9,.1))
print(Plot.SolarRad.AllSites.withcaption)
```

```{r, echo = FALSE}
#Faceted plot of mean monthly soil moisture at depth at each site
Plot.SoilMoist.Depth <-
  SCAN.CA.1519.summaries %>%
  rownames_to_column() %>%
  gather(key = key, value = value, soil_moist_perc_A_2in:soil_moist_perc_E_40in) %>%
  mutate(rowname = factor(rowname)) %>%
  ggplot(aes(date, value, color = key)) +
  geom_point(alpha = 0.85, cex = 2) +
  facet_wrap(vars(station_name_elev), nrow = 5, ) + #create facets by station
  xlab(expression(paste("Date"))) + 
  ylab(expression(paste("Mean Monthly Soil Moisture (%)"))) + 
  labs(title = "Soil Moisture at Depth Across California SCAN Stations", color = "Soil Depth") +
  scale_color_manual(values = c("#CCFFFF", "#66CCFF", "#3399CC", "#006699", "#003366"), labels = c("2 in", "4 in", "8 in",
                                "20 in", "40 in")) +
  scale_x_date(date_breaks = "1 year", date_labels = c("2020", "2015", "2016", "2017", "2018", "2019")) 
plot(Plot.SoilMoist.Depth)

#Create caption for plot
Caption.SoilMoist.Depth <- 
  ggplot() +
  theme(axis.line = element_blank(), axis.text =
          element_text(color = "black", size = 9),
          plot.title = element_text(size = 9, 
                                  color = "black")) + 
  labs(title = "Figure 2. Mean monthly soil moisture (%) at depth (inches) accross five Soil Climate Analysis Network (SCAN) stations in California from 2015-2019. Soil \n moisture was measured at soil depths of 2in, 4in, 8in, 20in, and 40in from 2015-2019. Each of the five SCAN stations are located at different elevations and in \n various regions of California. The value listed beside each station name denotes the elevation, in feet above sea level, of that given site. Please note: Data was \n not collected at the French Gulch station from May 2016-April 2017. However, this site was included in this analysis, as the trends evident in the present data are \n valuable to understanding trends at similar elevations/sites over time.")

#Combine plot and caption into single figure
Plot.SoilMoist.Depth.withcaption <-
  cowplot::plot_grid(Plot.SoilMoist.Depth, Caption.SoilMoist.Depth, 
                     align = 'h', axis = "a", 
                     nrow = 2, rel_heights = c(.9,.1))
print(Plot.SoilMoist.Depth.withcaption)
```

```{r, echo = FALSE}
#Faceted plot of mean monthly soil temperature at depth at each site
Plot.SoilTemp.Depth <-
  SCAN.CA.1519.summaries %>%
  rownames_to_column() %>%
  gather(key = key, value = value, soil_temp_A_2in:soil_temp_E_40in) %>%
  mutate(rowname = factor(rowname)) %>%
  ggplot(aes(date, value, color = key)) +
  geom_point(alpha = 0.85, cex = 2) +
  facet_wrap(vars(station_name_elev), nrow = 5, ) + #create facets by station
  xlab(expression(paste("Date"))) + 
  ylab(expression(paste("Mean Monthly Soil Temperature ("*~degree*F*")"))) + 
  labs(title = "Soil Temperature at Depth Across California SCAN Stations", color = "Soil Depth") +
  scale_color_manual(values = c("#CCFFFF", "#66CCFF", "#3399CC", "#006699", "#003366"), labels = c("2 in", "4 in", "8 in",
                                "20 in", "40 in")) +
  scale_x_date(date_breaks = "1 year", date_labels = c("2020", "2015", "2016", "2017", "2018", "2019")) 
plot(Plot.SoilTemp.Depth)

#Create caption for plot
Caption.SoilTemp.Depth <- 
  ggplot() +
  theme(axis.line = element_blank(), axis.text =
          element_text(color = "black", size = 9),
          plot.title = element_text(size = 9, 
                                  color = "black")) + 
  labs(title = "Figure 3. Mean monthly soil temperature (degrees Fahrenheit) at depth (inches) accross five Soil Climate Analysis Network (SCAN) stations in California from \n 2015-2019. Soil temperature was measured at soil depths of 2in, 4in, 8in, 20in, and 40in from 2015-2019. Each of the five SCAN stations are located at different \n elevations and in various regions of California. The value listed beside each station name denotes the elevation, in feet above sea level, of that given site. \n Please note: Data was not collected at the French Gulch station from May 2016-April 2017. However, this site was included in this analysis, as the trends evident \n in the present data are valuable to understanding trends at similar elevations/sites over time.")

#Combine plot and caption into single figure
Plot.SoilTemp.Depth.withcaption <-
  cowplot::plot_grid(Plot.SoilTemp.Depth, Caption.SoilTemp.Depth, 
                     align = 'h', axis = "a", 
                     nrow = 2, rel_heights = c(.9,.1))
print(Plot.SoilTemp.Depth.withcaption)
```

```{r, echo = FALSE}
#Faceted plot of mean monthly soil moisture by solar radiation at each site
Plot.SoilMoist.SolarRad <-
  SCAN.CA.1519.summaries %>%
  rownames_to_column() %>%
  gather(key = key, value = value, soil_moist_perc_A_2in:soil_moist_perc_E_40in) %>%
  mutate(rowname = factor(rowname)) %>%
  ggplot(aes(solar_rad_avg, value, color = key)) +
  geom_point(alpha = 0.85, cex = 2) +
  facet_wrap(vars(station_name_elev), nrow = 5, ) + #create facets by station
  xlab(expression(paste("Mean Monthly Solar Radiaion (W/m"^2*")"))) + 
  ylab(expression(paste("Mean Monthly Soil Moisture (%)"))) + 
  labs(title = "Soil Moisture by Solar Radiation Across California SCAN Stations", color = "Soil Depth") +
  scale_color_manual(values = c("#CCFFFF", "#66CCFF", "#3399CC", "#006699", "#003366"), labels = c("2 in", "4 in", "8 in",
                                "20 in", "40 in"))
plot(Plot.SoilMoist.SolarRad)

#Create caption for plot
Caption.SoilMoist.SolarRad <- 
  ggplot() +
  theme(axis.line = element_blank(), axis.text =
          element_text(color = "black", size = 9),
          plot.title = element_text(size = 9, 
                                  color = "black")) + 
  labs(title = "Figure 4. Mean monthly soil moisture (%) by mean monthly solar radiation (watts/square meter) accross five Soil Climate Analysis Network (SCAN) stations in \n California from 2015-2019. Soil moisture was measured at soil depths of 2in, 4in, 8in, 20in, and 40in from 2015-2019. Each of the five SCAN stations are located \n at different elevations and in various regions of California. The value listed beside each station name denotes the elevation, in feet above sea level, of that \n given site. Please note: Data was not collected at the French Gulch station from May 2016-April 2017. However, this site was included in this analysis, as the \n trends evident in the present data are valuable to understanding trends at similar elevations/sites over time.")

#Combine plot and caption into single figure
Plot.SoilMoist.SolarRad.withcaption <-
  cowplot::plot_grid(Plot.SoilMoist.SolarRad, Caption.SoilMoist.SolarRad,
                     align = 'h', axis = "a", 
                     nrow = 2, rel_heights = c(.9,.1))
print(Plot.SoilMoist.SolarRad.withcaption)
```

```{r, echo = FALSE}
#Faceted plot of mean monthly soil temperature by solar radiation at each site
Plot.SoilTemp.SolarRad <-
  SCAN.CA.1519.summaries %>%
  rownames_to_column() %>%
  gather(key = key, value = value, soil_temp_A_2in:soil_temp_E_40in) %>%
  mutate(rowname = factor(rowname)) %>%
  ggplot(aes(solar_rad_avg, value, color = key)) +
  geom_point(alpha = 0.85, cex = 2) +
  facet_wrap(vars(station_name_elev), nrow = 5, ) + #create facets by station
  xlab(expression(paste("Mean Monthly Solar Radiaion (W/m"^2*")"))) + 
  ylab(expression(paste("Mean Monthly Soil Temperature ("*~degree*F*")"))) +
  labs(title = "Soil Temperature by Solar Radiation Across California SCAN Stations", color = "Soil Depth") +
  scale_color_manual(values = c("#CCFFFF", "#66CCFF", "#3399CC", "#006699", "#003366"), labels = c("2 in", "4 in", "8 in",
                                "20 in", "40 in"))
plot(Plot.SoilTemp.SolarRad)

#Create caption for plot
Caption.SoilTemp.SolarRad <- 
  ggplot() +
  theme(axis.line = element_blank(), axis.text =
          element_text(color = "black", size = 9),
          plot.title = element_text(size = 9, 
                                  color = "black")) + 
  labs(title = "Figure 5. Mean monthly soil temperature (degrees Fahrenheit) by mean monthly solar radiation (watts/square meter) accross five Soil Climate Analysis \n Network (SCAN) stations in California from 2015-2019. Soil temperature was measured at soil depths of 2in, 4in, 8in, 20in, and 40in from 2015-2019. Each of \n the five SCAN stations are located at different elevations and in various regions of California. The value listed beside each station name denotes the \n elevation, in feet above sea level, of that given site. Please note: Data was not collected at the French Gulch station from May 2016-April 2017. However, \n this site was included in this analysis, as the trends evident in the present data are valuable to understanding trends at similar elevations/sites over time.")

#Combine plot and caption into single figure
Plot.SoilTemp.SolarRad.withcaption <-
  cowplot::plot_grid(Plot.SoilTemp.SolarRad, Caption.SoilTemp.SolarRad, 
                     align = 'h', axis = "a", 
                     nrow = 2, rel_heights = c(.9,.1))
print(Plot.SoilTemp.SolarRad.withcaption)
```

## Question 1: <insert specific question here and add additional subsections for additional questions below, if needed>

## Question 2: 




\newpage

# Summary and Conclusions


\newpage

# References
<add references here if relevant, otherwise delete this section> 
